---
title: "exam2"
output: word_document
---

# Clearing the R environment
``` {r}
rm(list=ls(all=TRUE))
```


# Working with College_Scorecard Data Sheet
## Loading the college_scorecarddataset in R, naming it  “college_scorecard”
```{r}
library(readxl)
college_scorecard <-
  read_excel("2021_exam2_data.xlsx", 4)
```

## Providing summary statistics for thecollege_scorecard dataset.
```{r}
summary(college_scorecard)
```

## Creating smaller dataset consisting only of data measured in 2014 and 2015 on formerstudents who graduated from four-year+ colleges and universities located inTexas (state_abbr: “TX”) and Louisiana (state_abbr: “LA”). 

``` {r}
library(dplyr)
library(tidyverse)
small_scorecard1 <- 
  subset(college_scorecard, year == c(2014, 2015))

#small_scorecard <-   
 # subset(small_scorecard1, state.abb == "LA|TX")
  ##^^ I could not figure out how to subset for the two states, kept receiving the error "Input has size 6018 but subscript `r` has size 50"
```





# Working with Acovados Data Sheet

## Loading the college_scorecarddataset in R, naming it  “college_scorecard”
```{r}
library(readxl)
avocados <-
  read_excel("2021_exam2_data.xlsx", 2)
```

## Creating new variable called year with the year avocadoes were sold
``` {r}
avocados <- 
  avocados %>% 
  mutate(year = lubridate::year(avocados$date))
```

## Deflating average price variable using WDI data
```{r}
library(WDI)

#creating deflator set
deflator_data =WDI(country = "all", indicator =c("NY.GDP.DEFL.ZS"),
                   start = 1960, # start of foreign aid data
                   end = 2019, # end of of foreign aid data
                   extra = FALSE, cache = NULL)

#renaming deflator variable
library(data.table)
setnames(deflator_data,"NY.GDP.DEFL.ZS", "deflator")

#selectin only US data
usd_deflator =subset(deflator_data, country=="United States")

#getting base year
subset(usd_deflator, deflator == 100)

# drop unnecessary variable
usd_deflator$country <- NULL
usd_deflator$iso2c <- NULL

#merging data
deflated_data =
  left_join(avocados,
            usd_deflator,
            by=c("year"))

#actually deflating the data
deflated_data$deflated_amount = deflated_data$average_price/(deflated_data$deflator/100)

#renamed data with deflated price
collapsed_avocados <- deflated_data


#cleaning workspace
rm(deflated_data)
rm(deflator_data)
rm(usd_deflator)

#showing output
head(collapsed_avocados)

#pivotting data
wide_avocados <-
  collapsed_avocados %>%
  pivot_wider(id_cols =c("year"),# unique IDs
              names_from = "date",# names for new wide vars
              values_from = "deflated_amount")# data to put in new wide vars

head(wide_avocados)
```

# Training Dataset Questions
``` {r}
#loading training dataset
library(readxl)
training <-
  read_excel("2021_exam2_data.xlsx", 3)
```


# Titanic Dataset Questions
``` {r}

library(readxl)
titanic <-
  read_excel("2021_exam2_data.xlsx", 1)

#providing summary statistics
summary(titanic)

##creating cross tab of survivorship by gender
male_passengers <- 
  subset(titanic, female == 0)

female_passengers <- 
  subset(titanic, female == 1)

#creating gender variable
titanic$gender = NA
titanic$gender[titanic$female ==1] = "Female"
titanic$gender[titanic$female ==0] = "Male"

#$checking table to esnure it all went through
table(titanic$gender, exclude = TRUE)

library(doBy)
summaryBy(survived ~ gender, data = titanic, FUN=c(mean,length))

titanic$first_class = NA
#ifelse(test = class==1, yes = )
```

#Ran out of time, spent too much on the first set of Qs

##Thanks Professors Findley and Denly!
